Clustering in a hyperbolic model of complex networks.

### 2019-03-18 concentration_argument.tex
Updates from Overleaf

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n}$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition, define for any constant $C > 0$,
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right).
\] 
Then the following holds.
\begin{enumerate}
\item For any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, 
\begin{equation}\label{eq:error_bound_int_rho_not_K}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}},
\end{equation}
as $n \to \infty$.
\item If in addition, $h(a_n) \sim h(b_n)$ whenever $a_n \sim b_n$, as $n \to \infty$. Then,
\begin{equation}\label{eq:concentration_h_rho}
	\int_0^\infty h(y) \rho(y,k_n) \alpha e^{-\alpha y^\prime} \dd y^\prime \sim  
		2\alpha \xi^{2\alpha} h(2\log(k_n/\xi)) k_n^{-(2\alpha + 1)},
\end{equation}
as $n \to \infty$.
\end{enumerate}
\end{lemma}

@@

\paragraph{Proof of the first statement.}
Recall (see proof of Proposition~\ref{prop:asymp}) that $\rho(y,k_n)$, as a function of $y$, is strictly increasing on $[0,a_n^-]$ and strictly decreasing on $[a_n^+,\infty)$. Therefore, by our assumption on $h(y)$,
\begin{align*}
	&\hspace{-20pt}\int_{\R_+ \setminus (a_n^-, a_n^+)} h(y) \rho(y,k_n) 
		\alpha e^{-\alpha y} \dd y\\
    &= \bigO{1} \int_0^{a_n^-} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y} \dd y 
    	+ \bigO{1}\int_{a_n^+}^{\infty} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y}x \dd y \\
    &= \bigO{1} \int_0^{a_n^-} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y 
   		+ \bigO{1} \int_{a_n^+}^{\infty} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y\\
   	&\le \bigO{1}\rho(a_n^-,k_n)\int_0^{a_n^-} e^{-(\alpha-\beta) y} \, \dd y
   		+ \bigO{1} \rho(a_n^+,k_n) \int_{a_n^+}^{\infty} e^{-(\alpha-\beta) y} \, \dd y.
\end{align*}
Since $\alpha - \beta > 0$, we conclude that
\begin{equation}\label{eq:concentration_lemma_integral_bound}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y
	= \bigO{1} \left(\rho(a_n^-,k_n) + \rho(a_n^+,k_n)\right) 
\end{equation}

@@

Since 
\[
	\frac{\mu(a_n^+)}{k_n} = \frac{\lambda_n^{+}}{k_n} = 1 + \epsilon_n + C \frac{\kappa_n}{k_n} 
	= 1 + \epsilon_n + C \sqrt{\frac{(1+\epsilon_n)\log((1+\epsilon_n)k_n)}{k_n}},
\]
and $x - \log(1 + x) \sim x^2/2$ as $x \to 0$, we get 
\begin{align*}
	\rho(a_n^+,k_n) 
	&\le \sqrt{2\pi} k_n^{-1/2} 
		e^{-k_n\left(\epsilon_n + C \frac{\kappa_n}{k_n} - \log\left(1 + \epsilon_n + C \frac{\kappa_n}{k_n}\right)\right)}\\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n + C \kappa_n/k_n\right)^2}{2}} \\
	&= \bigO{k_n^{-(1+C^2)/2}},		\numberthis \label{eq:concentration_lemma_bound_an+}
\end{align*}
where for the last line we used that
\[
	-k_n \frac{\left(\epsilon_n + C \kappa_n/k_n\right)^2}{2} = -\frac{C^2}{2}\log(k_n) + \bigT{1}.
\]
A similar analysis as above yields
\begin{equation}\label{eq:concentration_lemma_bound_an-}
	\rho(a_n^-,k_n) \le \bigT{1} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n - C \kappa_n/k_n\right)^2}{2}} = \bigO{k_n^{-(1+C^2)/2}}.
\end{equation} 
Plugging~\eqref{eq:concentration_lemma_bound_an-} and~\eqref{eq:concentration_lemma_bound_an+}  into~\eqref{eq:concentration_lemma_integral_bound} yields the result.

>>>>

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $(k_n)_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n}$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition, define for any constant $C > 0$,
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right).
\] 
Then the following holds.
\begin{enumerate}
\item For any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, 
\begin{equation}\label{eq:error_bound_int_rho_not_K}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}},
\end{equation}
as $n \to \infty$.
\item If in addition, $h(a_n) \sim h(b_n)$ whenever $a_n \sim b_n$, as $n \to \infty$. Then,
\begin{equation}\label{eq:concentration_h_rho}
	\int_0^\infty h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y \sim  
		2\alpha \xi^{2\alpha} h(2\log(k_n/\xi)) k_n^{-(2\alpha + 1)},
\end{equation}
as $n \to \infty$.
\end{enumerate}
\end{lemma}

@@

\paragraph{Proof of the first statement.}
Recall (see proof of Proposition~\ref{prop:asymp}) that $\rho(y,k_n)$, as a function of $y$, is strictly increasing on $[0,a_n^-]$ and strictly decreasing on $[a_n^+,\infty)$. Therefore, by our assumption on $h(y)$,
\begin{align*}
	&\hspace{-20pt}\int_{\R_+ \setminus (a_n^-, a_n^+)} h(y) \rho(y,k_n) 
		\alpha e^{-\alpha y} \dd y\\
    &= \bigO{1} \int_0^{a_n^-} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y} \dd y 
    	+ \bigO{1}\int_{a_n^+}^{\infty} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y} \dd y \\
    &= \bigO{1} \int_0^{a_n^-} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y 
   		+ \bigO{1} \int_{a_n^+}^{\infty} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y\\
   	&\le \bigO{1}\rho(a_n^-,k_n)\int_0^{a_n^-} e^{-(\alpha-\beta) y} \, \dd y
   		+ \bigO{1} \rho(a_n^+,k_n) \int_{a_n^+}^{\infty} e^{-(\alpha-\beta) y} \, \dd y.
\end{align*}
Since $\alpha - \beta > 0$, we conclude that
\begin{equation}\label{eq:concentration_lemma_integral_bound}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y
	= \bigO{1} \left(\rho(a_n^-,k_n) + \rho(a_n^+,k_n)\right) 
\end{equation}

@@

Since 
\[
	\frac{\mu(a_n^+)}{k_n} = \frac{\lambda_n^{+}}{k_n} = 1 + \epsilon_n + C \frac{\kappa_n}{k_n} 
	= 1 + \epsilon_n + C \sqrt{\frac{(1+\epsilon_n)\log((1+\epsilon_n)k_n)}{k_n}},
\]
and as $x - \log(1 + x) \sim x^2/2$ as $x \to 0$, we get 
\begin{align*}
	\rho(a_n^+,k_n) 
	&\le \sqrt{2\pi} k_n^{-1/2} 
		e^{-k_n\left(\epsilon_n + C \frac{\kappa_n}{k_n} - \log\left(1 + \epsilon_n + C \frac{\kappa_n}{k_n}\right)\right)}\\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n + C \kappa_n/k_n\right)^2}{2}} \\
	&= \bigO{k_n^{-(1+C^2)/2}},		\numberthis \label{eq:concentration_lemma_bound_an+}
\end{align*}
where for the last line we used that
\[
	-k_n \frac{\left(\epsilon_n + C \kappa_n/k_n\right)^2}{2} = -\frac{C^2}{2}\log(k_n) + \bigT{1}.
\]
A similar analysis as above yields
\begin{equation}\label{eq:concentration_lemma_bound_an-}
	\rho(a_n^-,k_n) \le \bigT{1} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n - C \kappa_n/k_n\right)^2}{2}} = \bigO{k_n^{-(1+C^2)/2}}.
\end{equation} 
Plugging~\eqref{eq:concentration_lemma_bound_an-} and~\eqref{eq:concentration_lemma_bound_an+}  into~\eqref{eq:concentration_lemma_integral_bound} yields the result.
